Learning constraints in tabular data

### 2016-05-26 Paper/ECAI/tabular.tex
Approach tweaks

  \begin{subfigure}{.70\textwidth}
  \begin{center}
    \includegraphics[width=1\textwidth]{figures/Demo2.png}
  \end{center}
  \vspace{-10pt}
  \caption{Example spreadsheet (black words and numbers only). Green borders indicate headerless tables; the blue elements are added to clarify some constraints and relations. \tias{TODO: headers not in green boxes} \sergey{show blocks? orientation only over blocks, not over tables?}\tias{Use specific constraint names, e.g. 'SUM-row' above Total in T1}}
  \label{fig:main_example}
\end{subfigure}
\hfill
\begin{subfigure}{.30\textwidth}
  {\footnotesize
    \begin{align*}
      % Table 1
      &~\ecseries{\range{T_{1}}{\rangeall}{1}} \\
%
      &~\ecperm{\range{T_{1}}{\rangeall}{1}}, \ecperm{\range{T_{1}}{\rangeall}{8}} \\
%
      &~\ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{3}}, \ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{4}} \\
      &~\ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{7}}, \ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{5}} \\
      &~\ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{6}}, \ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{10}} \\
%
      &~\ecsumr{\range{T_{1}}{\rangeall}{7}}{\range{T_{1}}{\rangeall}{\rangeto{3}{6}}} \\
%
      &~\ecaggif{\textit{SUM}}{\range{T_{1}}{\rangeall}{10}}{\range{T_{5}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{5}}{\rangeall}{2}} \\
%
      &~\ecaggif{\textit{MAX}}{\range{T_{1}}{\rangeall}{11}}{\range{T_{5}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{5}}{\rangeall}{2}} \\
%
	  % Table 2
      &~\ecsumc{\range{T_{2}}{1}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
%
      &~\ecaggc{\textit{AVERAGE}}{\range{T_{2}}{2}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
%
      &~\ecaggc{\textit{MAX}}{\range{T_{2}}{3}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}},  \\
%
      &~\ecaggc{\textit{MIN}}{\range{T_{2}}{4}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
%
	  % Table 4
      &~\eclookup{\range{T_{4}}{\rangeall}{2}}{\range{T_{4}}{\rangeall}{3}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{1}} \\
%
      &~\eclookup{\range{T_{4}}{\rangeall}{3}}{\range{T_{4}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}} \\
%
	  % Table 6
      &~\ecsumc{\range{T_{6}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{\rangeto{3}{6}}} \\
%
      &~\range{T_{6}}{\rangeall}{4} = PREV(\range{T_{6}}{\rangeall}{4}) + \range{T_{6}}{\rangeall}{2} - \range{T_{6}}{\rangeall}{3}
    \end{align*}}
  \vspace{-10pt}
  \caption{Constraints present in the running example (left), except \textit{ALLDIFFERENT} (16) and \textit{FOREIGNKEY} (5).}
  \label{fig:sol_example}
\end{subfigure}
\end{figure*}

@@

\subsubsection{Generating superblock assignments}
%The goal of this step is to prune out impossible groups from the beginning by looking at their properties rather than their actual content (\CSignature vs \CFunction).
Given a constraint template~$s$ and the set of all blocks~\blocks, the goal is to find all combinations of blocks that satisfy the constraint signature. Let there be $b$ blocks and let~$s$ have~$n$ arguments, then a simple generate-and-test method would have to test $b^n$ combinations (given that blocks can be repeated), which can grow very rapidly with $b$, depending on $n$.

@@

\subsubsection{Constraints over satisfying subgroups}
\label{sec:algo:subgr}
In the previous step superblock assignments are generated for a constraint template $s$ with $n$ arguments.
In this step, given such a superblock assignment $(B_1, \dots, B_n)$, the goal is to discover valid \textit{subassignments}, i.e. assignments of subblocks $(B'_1, \dots, B'_n)$ (for all $i$, $B'_i \subseteq B_i$) that satisfy both the signature and the definition of template~$s$.
While the previous step reasons about the \textit{properties} of blocks (type, size, length, ...), this step reasons about the actual \textit{content} of the vectors in the blocks, that is, the data.

@@

The search for subassignments is represented as the generic procedure $\findassignment$.
However, the actual implementation is specified for every constraint template separately, allowing different methods to be used based on the type of template.
One possible method is to reformulate the problem as a CSP, similar to the approach we employed in the last section.
Often, however, a simple generate-and-test approach will suffice.
In contrast to the previous step, the number of subblocks to generate here are typically manageable: $O(n*m^2)$ with $m$ the size of the largest block and $n$ the number of arguments of the template.
We only consider continuous ranges as subblocks, so for a block of size $m$ there are $m*(m+1)/2$ possible subblocks to consider.
Furthermore, few constraint solvers support floating point numbers, which are prevalent in spreadsheet.

@@

\begin{algorithm}[tbh]
  \begin{algorithmic}
    \footnotesize
    \Procedure{Subassignments}{$s$, $(B_1, \dots, B_n)$}
    \State $n \gets$ number of arguments of template $s$
    \State $A_{sub} \gets \emptyset$
    \ForAll{$(B_1', \dots, B_n') \in \{ (B_1', \dots, B_n') | \forall i: B_i' \sqsubseteq B_i \}$}
      \If{$disjoint(\{B_1', \dots, B_n'\}) \land Test((B_1',\dots,B_n'))$}
        \State $A_{sub} \gets A_{sub} \cup \{ (B_1', \dots, B_n') \}$
      \EndIf
    \EndFor
\Return $A_{sub}$
\EndProcedure
\end{algorithmic}
\caption{Generate-and-test for $\findassignment$}
\label{algo:gen-test-subblocks}
\end{algorithm}

@@

\subsubsection{Redundancy}
We consider two types of redundancy that we aim to eliminate during the search.
As noted in Section~\ref{sec:form:dependencies}, for some constraint templates there are \textit{symmetries} over the arguments that lead to trivially equivalent solutions, for example, $\ecprod{B_R'}{B_1'}{B_2'} \Leftrightarrow \ecprod{B_R'}{B_2'}{B_1'}$.
Such duplicates are avoided by defining a \textit{canonical} form for these constraints.
In practice, we define an arbitrary but fixed block ordering and require those blocks that are interchangeable to adhere to this ordering.
\samuel{Mention B1*B2 vs PRODUCT?}

@@

Note the the template $\ecdiff{B_r}{B_1}{B_2}$ can be rewritten as $B_1 = B_r + B_2$.
Our choice to use difference is a hardcoded preference based on the prevalence of this type of constraint in financial spreadsheet data.

>>>>

  \begin{subfigure}{.70\textwidth}
  \begin{center}
    \includegraphics[width=1\textwidth]{figures/Demo2.png}
  \end{center}
  \vspace{-10pt}
  \caption{Example spreadsheet (black words and numbers only). Green borders indicate headerless tables; the blue elements are added to clarify some constraints and relations. \tias{Use specific constraint names, e.g. 'SUM-row' above Total in T1}}
  \label{fig:main_example}
\end{subfigure}
\hfill
\begin{subfigure}{.30\textwidth}
  {\footnotesize
    \begin{align*}
      % Table 1
      &~\ecseries{\range{T_{1}}{\rangeall}{1}} \\
%
      &~\ecperm{\range{T_{1}}{\rangeall}{1}}, \ecperm{\range{T_{1}}{\rangeall}{8}} \\
%
      &~\ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{3}}, \ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{4}} \\
      &~\ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{7}}, \ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{5}} \\
      &~\ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{6}}, \ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{10}} \\
%
      &~\ecsumr{\range{T_{1}}{\rangeall}{7}}{\range{T_{1}}{\rangeall}{\rangeto{3}{6}}} \\
%
      &~\ecaggif{\textit{SUM}}{\range{T_{1}}{\rangeall}{10}}{\range{T_{5}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{5}}{\rangeall}{2}} \\
%
      &~\ecaggif{\textit{MAX}}{\range{T_{1}}{\rangeall}{11}}{\range{T_{5}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{5}}{\rangeall}{2}} \\
%
	  % Table 2
      &~\ecsumc{\range{T_{2}}{1}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
%
      &~\ecaggc{\textit{AVERAGE}}{\range{T_{2}}{2}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
%
      &~\ecaggc{\textit{MAX}}{\range{T_{2}}{3}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}},  \\
%
      &~\ecaggc{\textit{MIN}}{\range{T_{2}}{4}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
%
	  % Table 4
      &~\eclookup{\range{T_{4}}{\rangeall}{2}}{\range{T_{4}}{\rangeall}{3}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{1}} \\
%
      &~\eclookup{\range{T_{4}}{\rangeall}{3}}{\range{T_{4}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}} \\
%
	  % Table 6
      &~\ecsumc{\range{T_{6}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{\rangeto{3}{6}}} \\
%
      &~\range{T_{6}}{\rangeall}{4} = PREV(\range{T_{6}}{\rangeall}{4}) + \range{T_{6}}{\rangeall}{2} - \range{T_{6}}{\rangeall}{3}
    \end{align*}}
  \vspace{-10pt}
  \caption{Constraints present in the running example (left), except \textit{ALLDIFFERENT} (16) and \textit{FOREIGNKEY} (5).\samuel{Not the desired constraints ask : \tias{}}}
  \label{fig:sol_example}
\end{subfigure}
\end{figure*}

@@

\subsubsection{Generating superblock assignments}
Given a constraint template~$s$ and the set of all blocks~\blocks, the goal is to find all combinations of blocks that satisfy the constraint signature. Let there be $b$ blocks and let~$s$ have~$n$ arguments, then a simple generate-and-test method would have to test $b^n$ combinations (given that blocks can be repeated), which can grow very rapidly with $b$, depending on $n$.

@@

\subsubsection{Constraints over satisfying subgroups}
\label{sec:algo:subgr}
In the previous step superblock assignments are generated for a constraint template $s$ with $n$ arguments.
In this step, given such a superblock assignment $(B_1, \dots, B_n)$, the goal is to discover valid \textit{subassignments}, i.e. assignments of subblocks $(B'_1, \dots, B'_n)$ (for all $i$, $B'_i \subseteq B_i$) that satisfy both the signature and the definition of template~$s$.
\samuel{Is this repetition?}
While the previous step reasons about the \textit{properties} of blocks (type, size, length, ...), this step reasons about the actual \textit{content} of the vectors in the blocks, that is, the data.

@@

The search for subassignments is represented as the generic procedure $\findassignment$.
However, the actual implementation is specified for every constraint template separately, allowing different methods to be used based on the type of template.
One possible method is to reformulate the problem as a CSP, similar to the approach we employed in the last section.
Often, however, a simple generate-and-test approach will suffice.
In contrast to the previous step, the number of subblocks to generate here are typically manageable: $O(n*m^2)$ with $m$ the size of the largest block and $n$ the number of arguments of the template.
We only consider continuous ranges as subblocks, so for a block of size $m$ there are $m*(m+1)/2$ possible subblocks to consider.
Furthermore, few constraint solvers support floating point numbers, which are prevalent in spreadsheets.

@@

\begin{algorithm}[tbh]
  \begin{algorithmic}
    \footnotesize
    \Procedure{Subassignments}{$s$, $(B_1, \dots, B_n)$}
    \State $n \gets$ number of arguments of template $s$
    \State $A_{sub} \gets \emptyset$
    \ForAll{$(B_1', \dots, B_n') \in \{ (B_1', \dots, B_n') | \forall i: B_i' \sqsubseteq B_i \}$}
      \If{$disjoint(\{ B_1', \dots, B_n' \}) \land Test((B_1',\dots,B_n'))$}
        \State $A_{sub} \gets A_{sub} \cup \{ (B_1', \dots, B_n') \}$
      \EndIf
    \EndFor
\Return $A_{sub}$
\EndProcedure
\end{algorithmic}
\caption{Generate-and-test for $\findassignment$}
\label{algo:gen-test-subblocks}
\end{algorithm}

@@

\subsubsection{Redundancy}
We consider two types of redundancy that we aim to eliminate during the search.
As noted in Section~\ref{sec:form:dependencies}, for some constraint templates there are \textit{symmetries} over the arguments that lead to trivially equivalent solutions, for example, $\ecprod{B_R'}{B_1'}{B_2'} \Leftrightarrow \ecprod{B_R'}{B_2'}{B_1'}$.
Such duplicates are avoided by defining a \textit{canonical} form for these constraints.
In practice, we define an arbitrary but fixed block ordering and require those blocks that are interchangeable to adhere to this ordering.
Moreover, \sname currently uses product (\ecprod{B_R'}{B_1'}{B_2'}) instead of division (\eccalc{B_1'}{B_R' / B_2'}) and difference (\ecdiff{B_r}{B_1}{B_2}) instead of sum (\eccalc{B_1}{B_r + B_2}).
These choices are hardcoded preferences based on the prevalence of difference and product in financial spreadsheet data, however, these constraints can be trivially rewritten if necessary.

@@
